Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-08-28 manuscript/manuscript.tex
small appendix updates to match ms
small edits from PBA on appendices

Forecasting future states of populations has taken on new urgency as the
rate of climate change increases. Traditional plant population models,
typically built using detailed demographic observations from one study
site, can address climate change impacts at one location, but are
difficult to scale up to the landscape and regional scales relevant to
management decisions. One proposal to overcome the limitations of
traditional population models is to build models using population-level
data that is much easier to collect over broad spatial scales than
individual-level data. However, such models ignore the fact that climate
drives population growth through its influence on individual
performance. Here, we test the consequences of aggregating individual
responses when forecasting climate change impacts on four perennial
grass species in a semi-arid Montana grassland. We parameterized two
population models, one based on individual-level data (survival, growth
and recruitment) and one on population-level data from the same dataset
(percent cover of quadrats), and compared their accuracy, precision, and
sensitivity to climate. The individual level model was more accurate and
precise than the aggregate level model when predicting out-of-sample
observations. When comparing climate effects from both models, the
population-level model tends to miss important climate effects from at
least one vital rate for each species. Increasing the sample size at the
population-level would not necessarily reduce forecast uncertainty; the
way to reduce uncertainty is to capture unique climate dependence of
individual vital rates. Our analysis indicates that there is no shortcut
to forecasting climate change impacts on plant populations --- detailed
demographic data is essential. Despite the superiority of the
individual-level model, the forecasts it generated still were too
uncertain to be useful for decision-makers. We need new methods to
collect demographic data more efficiently across environmental gradients
in space and time.

@@

Perhaps the greatest challenge for ecology in the 21st century is to
forecast the impacts of environmental change (Clark et al. 2001, Petchey
et al. 2015). Forecasts require sophisticated modeling approaches that
fully account for uncertainty and variability in both ecological process
and model parameters (Luo et al. 2011, but see Perretti et al. 2013 for
an argument against modeling the ecological process). This requires
large amounts of data collected over broad spatio-temporal extents, and
state-of-the-science modeling techniques cannot overcome data
limitations. Population models are a case in point.

Population models are important tools for predicting the impacts of
environmental change on species persistence and abundance. But
reconciling the scales at which population models are parameterized and
the scales at which environmental changes play out remains a challenge
(Clark et al. 2010, 2012, Freckleton et al. 2011, Queenborough et al.
2011). The problem is that most population models are built using data
from a single study site because collecting those data, which involves
tracking the fates of individuals plants, is so difficult. The resulting
models cannot be applied to the landscape and regional scales relevant
to decision-makers without information about how the fitted parameters
respond to spatial variation in biotic and abiotic drivers (Sæther et
al. 2007). The limited spatial extent of individual-level demographic
datasets constrains our ability to use population models to address
applied questions about the consequences of climate change.

The inability of most population models to address landscape-scale
problems may explain why land managers and conservation planners have
embraced species distribution models (SDMs) (see Guisan and Thuiller
2005 for a review). SDMs typically rely on easy-to-collect
presence/absence data (but see Clark et al. 2014 for new methods) and
remotely-sensed environmental covariates that allow researchers to model
large spatial extents (e.g., Maiorano et al. 2013). Thus, it is
relatively straightforward to parameterize and project SDMs over
landscapes and regions, the scales at which many land-use decisions are
made. However, the limitations of SDMs are well known (Pearson and
Dawson 2003, Elith and Leathwick 2009, Araújo and Peterson 2012).
Ideally, researchers would provide managers with landscape-scale
population models, combining the extent of SDMs with information about
dynamics and species abundances (Schurr et al. 2012, Merow et al. 2014).

Aggregate measures of population status, rather than individual
performance, offer an intriguing alternative for modeling populations
(Clark and Bjørnstad 2004, Freckleton et al. 2011). Population-level
data cannot provide inference about demographic mechanisms, but it might
be sufficient for modeling future population states, especially since it
is feasible to collect across broad spatial extents (e.g., Queenborough
et al. 2011). The choice between individual and population-level data
involves a difficult trade-off: while individual-level data leads to
more mechanistic models, population-level data leads to models that can
be applied over greater spatial and temporal extents. An open question
is how much forecasting skill is lost when we build models based on
population rather than individual-level data.

To date, most empirical population modelers have relied on
individual-level data, with few attempts to capitalize on
population-level measures. An important exception was an effort by
Taylor and Hastings (2004) to model the population growth rate of an
invasive species to identify the best strategies for invasion control.
They used a ``density-structured'' model where the state variable is a
discrete density state rather than a continuous density measure. Such
models do not require individual level demographic data and can
adequately describe population dynamics.

Building on Taylor and Hastings (2004), Freckleton et al. (2011) showed
that density-structured models compare well to continuous models in
theory, and Queenborough et al. (2011) demonstrated the application of
such methods in a study on arable weeds. In particular, Queenborough et
al. (2011) provide empirical evidence that density-structured models are
capable of reproducing population dynamics, even if some precision is
lost when compared to fully continuous models. The study by Queenborough
et al. (2011) included data from 500 fields (4 hectares each) in 49
farms, all collected by two people in 6 weeks. This is far more data
from a far greater spatial extent than would be possible if measuring
individual plant demography (in a world of limited time and money). The
appeal of density-structured approaches is clear. However, none of these
models included environmental covariates.

Addressing climate change questions with models fit to population-level
data is potentially problematic. It is individuals, not populations,
that respond to climate variables (Clark et al. 2012). Ignoring this
fact amounts to an ``ecological fallacy'', where inference about the
individual relies on statistical inference on the group (Piantadosi et
al. 1988). Population growth (or decline) is the outcome of demographic
processes like survival, growth, and recruitment that occur at the level
of individual plants. Climate can affect each demographic process in
unique, potentially opossing, ways (Dalgleish et al. 2011). These unique
climate responses may be difficult to resolve in statistical models
based on population-level data where demographic processes are not
identifiable. If important climate drivers are missed because they are
averaged over in population-level data, then population models built
with such data will make uninformative or unreliable forecasts when
confronted with novel climate covariates.

Here, we ask whether statistical and population models based on
aggregated, population-level data can predict climate change impacts as
well as models based on individual-level data. We used a unique
demographic dataset that tracks the fates of individual plants from four
species over 14 years to build two kinds of single-species population
models, traditional models using individual growth, survival, and
recruitment data and alternative models based on basal cover. In both
models, interannual variation is explained, in part, by climate
covariates. We then performed simulations to quantify the sensitivities
of species' cover to small perturbations in average precipitation and
temperature. Individual level models were clearly better, but,
unfortunately, still too uncertain to inform management decisions.

@@

In this paper, we model populations based on two levels of data:
individual and quadrat (Fig. 2). The individual data is the ``raw''
data. For the quadrat level we data we simply sum individual basal cover
for each quadrat by species. This is equivalent to a near-perfect census
of quadrat percent cover because previous analysis shows that
measurement error at the individual level is small (Chu and Adler 2014).
Based on these two datasets we can compare population models built using
individual level data and aggregated quadrat level data.

@@

At both levels of inference (individual and quadrat), the building
blocks of our population models are vital rate regressions. For
individual-level data, we fit models for survival, growth, and
recruitment for each species. At the quadrat-level, we fit a single
regression model for population growth. We describe the statistical
models separately since fitting the models required different
approaches. All models contain five climate covariates that we chose
\emph{a priori}: ``water year'' precipitation at \emph{t}-1 (lagppt);
April through June precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and
ppt2, respectively) and April through June temperature at \emph{t}-1 and
\emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is
the observation year. We also include interactions among same-year
climate covariates (e.g., ppt1 $\times$ TmeansSpr1) and climate $\times$
size interactions. Climate $\times$ size interactions are for climate
main effects only; we do not include interactions between size and pairs
of interacting climate effects.

We fit all models using a hierarchical Bayesian approach. The models are
fully descibed in Appendix A, so here we focus on the main process and
the model likelihood. For the likelihood models, $\textbf{y}^X$ is
always the relevant vector of observations for vital rate \emph{X}
($X = S, G, R, or P$ for survival, growth, recruitment, or population
growth). For example, $\textbf{y}^S$ is a vector of 0s and 1s indicating
whether a genet survives from \emph{t} to \emph{t+1}, or not.

@@

Our data allows us to track new recruits, but we cannot assign a
specific parent to new genets. So, for recruitment, we work at the
quadrat level and model the number of new individuals of species $j$ in
quadrat $q$ recruiting at time $t+1$ as a function of quadrat
``effective cover'' ($A'$) in the previous year ($t$). Effective cover
is a mixture of observed cover ($A$) in the focal quadrat ($q$) and the
mean cover across the entire group ($\bar{A}$) of $Q$ quadrats in which
$q$ is located:

@@

The statistical approach used to model vital rates using aggregated data
depends on the type of data collected. In our case, and as is often the
case with census data, we have percent cover data (which can easily be
transformed to proportion data). We first considered fitting three vital
rate models analagous to those we fit at the individual level: one for
probability of extirpation within a quadrat (analagous to survival), one
for cover change within a quadrat (analagous to growth), and one for
probability of colonization within a quadrat (analagous to recruitment).
However, within-quadrat extirpation and colonization events were rare in
our time series ($N=9$ and $N=10$, respectively, across all species).
Given the broad spatial distribution of the quadrats we are studying, it
is safe to assume that these events are in fact rare enough to be
ignored for our purposes. So we constrained our statistical modeling of
vital rates at the population level to change in percent cover within
quadrats. For the remaining discussion of statistical modeling we refer
to proportion data, which is simply percent data divided by 100.

An obvious choice for fitting a linear model to proportion data is beta
regression because the support of the beta distribution is {[}0,1{]},
not including true zeros or ones. However, when we used fitted model
parameters from a beta regression in a quadrat-based population model,
the simulated population tended toward 100\% cover for all species. We
therefore chose a more constrained modeling approach based on a
truncated log-normal likelihood. The model for quadrat cover change
($G$) from time $t$ to $t+1$ is

@@

Our Bayesian approach to fitting the vital rate models required choosing
appropriate priors for unknown parameters and deciding which, if any, of
those priors should be hierarchical. We decided to fit models where all
terms were fit by species. Within a species, we fit yearly size effects
and yearly intercepts hierarchically where year-specific coefficients
were drawn from global distributions representing the mean size effect
and intercept. We used uninformative priors (Appendix A).

All of our analyses (model fitting and simulating) were conducted in R
(R Core Development Team 2013). We used the `No-U-Turn' MCMC sampler in
Stan (Stan Development Team 2014a) to estimate the posterior
distributions of model parameters using the package `rstan' (Stan
Development Team 2014b). We obtained posterior distributions for all
model parameters from three parallel MCMC chains run for 1,000
iterations after discarding an initial 1,000 iterations. We recognize
such short MCMC chains may surprise those more familiar with other MCMC
samplers (i.e.~JAGS or WinBUGS), but the Stan sampler is exceptionally
efficient, which reduces the number of iterations needed to achieve
convergence. We assessed convergence visually and made sure scale
reduction factors for all parameters were less than 1.01. For the
purposes of including parameter uncertainty in our population models, we
saved the final 1,000 iterations from each of the three MCMC chains to
be used as randomly drawn values during population simulation. This step
alleviates the need to reduce model parameters by model selection since
sampling from the full parameter space in the MCMC ensures that if a
parameter broadly overlaps zero, on average the effect in the population
models will also be near zero. We report the posterior mean, standard
deviation, and 95\% Bayesian Credible Intervals for every parameter of
each model for each species in Appendix B.

@@

With the posterior distribution of the vital rate statistical models in
hand, it is straightforward to simulate the population models. We used
an Integral Projection Model (IPM) to model populations based on
individual level data (Ellner and Rees 2006) and a quadrat-based version
of an individually-based model (Quadrat-Based Model, QBM) to model
populations based on quadrat level data. We describe each in turn.

@@

We use an environmentally stochastic IPM (Mark and Ellner 2009) that
includes the climate covariates from the vital rate statistical models.
In all simulations we ignore the random year effects so that interannual
variation is driven solely by climate. We fit the random year effects in
the vital rate regressions to avoid over-attributing variation to
climate covariates. Our IPM follows the specification of Chu and Adler
(2015) where the population of species \emph{j} is a density function
$n(u_{j},t)$ giving the density of sized-\emph{u} genets at time
\emph{t}. Genet size is on the natural log scale, so that $n(u_{j},t)du$
is the number of genets whose area (on the arithmetic scale) is between
$e^{u_{j}}$ and $e^{u_{j}+du}$. The density function for any size
\emph{v} at time $t+1$ is

\begin{equation}
n(v_{j},t+1) = \int_{L_{j}}^{U_{j}} k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}}(u_{j}))n(u_{j},t)
\end{equation}

where $k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}})$ is the population kernel
that describes all possible transitions from size $u$ to $v$ and
$\bar{\bold{w_{j}}}$ is a vector of estimates of average crowding
experienced from all other species by a genet of size $u_j$ and species
$j$. The integral is evaluated over all possible sizes between
predefined lower (\emph{L}) and upper (\emph{U}) size limits that extend
beyond the range of observed genet sizes.

Since the IPM is spatially-implicit, we cannot calculate neighborhood
crowding for specific genets ($w_{ij}$). Instead, we use an
approximation ($\bar{\bold{w_{j}}}$) that captures the essential
features of neighborhood interactions (Adler et al. 2010). We also use a
`no-overlap' rule for conspecific genets to approximate the
overdispersion of large genets in space (Adler et al. 2010, 2012, Chu
and Adler 2015).

@@

\begin{equation}
k_{j}(v_{j},u_{j},\bar{\bold{w_{j}}}) = S_j(u_j, \bar{\bold{w_{j}}}(u_{j}))G_j(v_{j},u_{j},\bar{\bold{w_{j}}}(u_{j})) + R_j(v_{j},u_{j},\bar{\bold{w_{j}}}),
\end{equation}

which means we are calculating growth (\emph{G}) for individuals that
survive (\emph{S}) from time \emph{t} to \emph{t+1} and adding in newly
recruited (\emph{R}) individuals of an average sized one-year-old genet
for the focal species. Our stastical model for recruitment (\emph{R},
described above) returns the number of new recruit produced per quadrat.
Following previous work (Adler et al. 2012, Chu and Adler 2015), we
assume that fecundity increases linearly with size
($R_j(v_{j},u_{j},\bar{\bold{w_{j}}}) = e^{u_j}R_j(v_{j},\bar{\bold{w_{j}}})$)
to incorporate the recruitment function in the spatially-implicit IPM.

We used random draws from the final 1,000 iterations from each of three
MCMC chains to introduce stochasticity into our population models. At
each time step, we randomly selected climate covariates from one of the
14 observed years. Then, we drew the full parameter set (climate effects
and density-dependence fixed effects) from a randomly selected MCMC
iteration. Using this approach, rather than simply using coefficient
point estimates, ensures that relatively unimportant climate covariates
(those that broadly overlap 0) have little effect on the simulation
results. Since our focus was on the contribution of climate covariates
to population states, we set the random year effects and the random
group effects to zero.

@@

Our quad-based model (QBM) perfectly mirrors its statistical description
(Eqs. 9-10). We use the same approach for drawing parameter values as
described for the IPM. After drawing the appropriate parameter set, we
calculate the mean response (population cover at \emph{t+1} = $x_{t+1}$)
according to Eq. 9. We then make a random draw from a {[}0,1{]}
truncated lognormal distribution with mean equal to $x_{t+1}$ from Eq. 9
and the variance estimate from the fitted model. We can then project the
model forward by drawing a new parameter set (unique to climate year and
MCMC iteration) at each timestep. As with the IPM, random year effects
are ignored for all simulations.

@@

To test each model's ability to forecast population state, we made
out-of-sample predictions using leave-one-year-out cross validation. For
both levels of modeling, we fit the vital rate models using observations
from all years except one, and then used those fitted parameters in the
population models to perform a one-step-ahead forecast for the year
whose observations were withheld from model fitting. Within each
observation year, several quadrats were sampled. So we made predictions
for each observed quadrat in the focal year, initializing each
simulation with cover in the quadrat the previous year. Since we were
making quadrat-specific predictions, we incorporated the group effect on
the intercept for both models. We repeated this procedure for all 13
observation years, making 100 one-step-ahead forecasts for each
quadrat-year combination with parameter uncertainty included via random
draw from the MCMC chain as described above. Random year effects were
set to zero since year effects cannot be assigned to unobserved years.

This model validation allowed us to compare accuracy and precision of
the two modeling approaches (IPM versus QBM). We first calculated the
median predicted cover across the 100 simulations for each quadrat-year
and then calculated the absolute error as the absolute value of the
difference between the observed cover for a given quadrat-year and the
median prediction. To arrive at mean absolute error (MAE), we then
averaged the absolute error within each species across the quadrat-year
specific errors. We use MAE as our measure of accuracy. To measure
precision we calculated the distance between the upper and lower 90th
quantiles of the 100 predictions and averaged this value over
quadrat-years for each species.

@@

Our expectation is that the IPM will produce more accurate and precise
forecasts. This could be due to (1) the smaller sample size of the
quadrat level data sets compared to the individual level data sets,
leading to larger parameter uncertainty for the QBM, or (2) the QBM
climate effects being weakly associated with one or more vital rate
climate effects at the individual level. To assess the impact of sample
size on QBM parameter uncertainty we refit the QBM statistical model
(Eqs. 9-10) after removing sets of 2, 5, 10, and 15 quadrats. We fit 10
models at each level of quadrat removal (2, 5, 10, 15 quadrats),
removing a different randomly selected set of quadrats for each fit. We
calculated the standard deviation of climate main effects (pptLag, ppt1,
ppt2, TmeanSpr1, and TmeanSpr2) for each model and averaged those over
replicates within each set of quadrat removals. This allowed us to
regress sample size against parameter uncertainty.

To see if the QBM climate effects are correlated, or not, with climate
effects for each vital rate model for the IPM, we simply regressed the
QBM climate coefficients against each vital rate model's climate
coefficients and calculated Pearson's $\rho$. Strong correlations
indicate the QBM is capable of detecting climate effects associated with
individual vital rates. A weak correlation indicates the QBM ``misses''
the climate effect on a particular vital rate.

@@

Mark, R., and S. P. Ellner. 2009. Integral projection models for
populations in temporally varying environments. Ecological Monographs
79:575--594.

>>>>

Rapid climate change has generated growing interest in forecasts of
future population trajectories. Traditional population models, typically
built using detailed demographic observations from one study site, can
address climate change impacts at one location, but are difficult to
scale up to the landscape and regional scales relevant to management
decisions. An alternative is to build models using population-level data
that is much easier to collect over broad spatial scales than
individual-level data. However, such models ignore the fact that climate
drives population growth through its influence on individual
performance. Here, we test the consequences of aggregating individual
responses when forecasting climate change impacts on four perennial
grass species in a semi-arid grassland in Montana, USA. We parameterized
two population models based on the same dataset, one based on
individual-level data (survival, growth and recruitment) and one on
population-level data (percent cover), and compared their accuracy,
precision, and sensitivity to climate variables. The individual-level
model was more accurate and precise than the aggregated model when
predicting out-of-sample observations. When comparing climate effects
from both models, the population-level model missed important climate
effects from at least one vital rate for each species. Increasing the
sample size at the population-level would not necessarily reduce
forecast uncertainty; the way to reduce uncertainty is to capture unique
climate dependence of individual vital rates. Our analysis indicates
that there is no shortcut to forecasting climate change impacts on plant
populations --- detailed demographic data is essential. Despite the
superiority of the individual-level model, the forecasts it generated
still were too uncertain to be useful for decision-makers. We need new
methods to collect demographic data efficiently across environmental
gradients in space and time.

@@

Perhaps the greatest challenge for ecology in the 21st century is to
forecast the impacts of environmental change (Clark et al. 2001, Petchey
et al. 2015). Forecasts require sophisticated modeling approaches that
fully account for uncertainty and variability in both ecological process
and model parameters (Luo et al. 2011, but see Perretti et al. 2013 for
an argument against modeling the ecological process). The increasing
statistical sophistication of population models (Rees and Ellner 2009)
makes them promising tools for predicting the impacts of environmental
change on species persistence and abundance. But reconciling the scales
at which population models are parameterized and the scales at which
environmental changes play out remains a challenge (Clark et al. 2010,
2012, Freckleton et al. 2011, Queenborough et al. 2011). The problem is
that most population models are built using data from a single study
site because collecting those data, which involves tracking the fates of
individuals plants, is so difficult. The resulting models cannot be
applied to the landscape and regional scales relevant to decision-making
without information about how the fitted parameters respond to spatial
variation in biotic and abiotic drivers (Sæther et al. 2007). The
limited spatial extent of individual-level demographic datasets
constrains our ability to use population models to address applied
questions about the consequences of climate change.

The inability of most population models to address landscape-scale
problems may explain why land managers and conservation planners have
embraced species distribution models (SDMs) (see Guisan and Thuiller
2005 for a review). SDMs typically rely on easy-to-collect
presence/absence data (but see Clark et al. 2014 for new methods) and
remotely-sensed environmental covariates that allow researchers to model
large spatial extents (e.g., Maiorano et al. 2013). Thus, it is
relatively straightforward to parameterize and project SDMs over
landscapes and regions. However, the limitations of SDMs are well known
(Pearson and Dawson 2003, Elith and Leathwick 2009, Araújo and Peterson
2012). Ideally, researchers would provide managers with landscape-scale
population models, combining the extent of SDMs with information about
dynamics and species abundances (Schurr et al. 2012, Merow et al. 2014).

Aggregate measures of population status, rather than individual
performance, offer an intriguing alternative for modeling populations
(Clark and Bjørnstad 2004, Freckleton et al. 2011). Population-level
data cannot provide inference about demographic mechanisms, but might be
sufficient for modeling future population states, especially since such
data are feasible to collect across broad spatial extents (e.g.,
Queenborough et al. 2011). The choice between individual and
population-level data involves a difficult trade-off: while
individual-level data leads to more mechanistic models, population-level
data leads to models that can be applied over greater spatial and
temporal extents. An open question is how much forecasting skill is lost
when we build models based on population rather than individual-level
data.

To date, most empirical population modelers have relied on
individual-level data, with few attempts to capitalize on
population-level measures. An important exception was an effort by
Taylor and Hastings (2004) to model the population growth rate of an
invasive species to identify the best strategies for invasion control.
They used a ``density-structured'' model where the state variable is a
discrete density state rather than a continuous density measure. Such
models do not require individual-level demographic data and can
adequately describe population dynamics. Building on Taylor and Hastings
(2004), Freckleton et al. (2011) showed that density-structured models
compare well to continuous models in theory, and Queenborough et al.
(2011) provide empirical evidence that density-structured models are
capable of reproducing population dynamics at landscape spatial scales,
even if some precision is lost when compared to fully continuous models.
The appeal of density-structured approaches is clear. However, none of
these models included environmental covariates.

Addressing climate change questions with models fit to population-level
data is potentially problematic. It is individuals, not populations,
that respond to climate variables (Clark et al. 2012). Ignoring this
fact amounts to an ``ecological fallacy'', where inference about the
individual relies on statistical inference on the group (Piantadosi et
al. 1988). Population growth (or decline) is the outcome of demographic
processes such as survival, growth, and recruitment that occur at the
level of individual plants. Climate can affect each demographic process
in unique, potentially opossing, ways (Dalgleish et al. 2011). These
unique climate responses may be difficult to resolve in statistical
models based on population-level data where demographic processes are
not identifiable. If important climate effects are missed because of the
aggregation inherent in in population-level data, then population models
built with such data will make uninformative or unreliable forecasts.

Here, we compare the forecasting skill of statistical and population
models based on aggregated, population-level data with models based on
individual-level data. We used a unique demographic dataset that tracks
the fates of individual plants from four species over 14 years to build
two kinds of single-species population models, traditional models using
individual growth, survival, and recruitment data and alternative models
based on basal cover. In both models, interannual variation is
explained, in part, by climate covariates. We first quantify forecasting
skill using cross-validation. We then performed simulations to quantify
the sensitivities of species' cover to small perturbations in average
precipitation and temperature. Based on the cross-validation results,
predictions of individual level models were clearly better, but,
unfortunately, still too uncertain to inform management decisions.

@@

We modeled each grass population based on two levels of data: individual
and quadrat (Fig. 2). The individual data is the ``raw'' data. For the
quadrat-level we data we simply sum individual basal cover for each
quadrat by species. This is equivalent to a near-perfect census of
quadrat percent cover because previous analysis shows that measurement
error at the individual-level is small (Chu and Adler 2014). Based on
these two datasets we can compare population models built using
individual-level data and aggregated, quadrat-level data.

@@

At both levels of inference (individual and quadrat), the building
blocks of our population models are vital rate regressions. For
individual-level data, we fit regressions for survival, growth, and
recruitment for each species. At the quadrat-level, we fit a single
regression model for population growth. We describe the statistical
models separately since fitting the models required different
approaches. All models contain five climate covariates that we chose
\emph{a priori}: ``water year'' precipitation at \emph{t}-1 (lagppt);
April through June precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and
ppt2, respectively) and April through June temperature at \emph{t}-1 and
\emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is
the observation year. We also include interactions among same-year
climate covariates (e.g., ppt1 $\times$ TmeansSpr1) and climate $\times$
size interactions. Climate $\times$ size interactions are for climate
main effects only; we do not include interactions between size and pairs
of interacting climate effects.

We fit all models using a hierarchical Bayesian approach. The models are
fully descibed in Appendix A, so here we focus on the main process and
the model likelihood. For the likelihood models, $\textbf{y}^X$ is
always the relevant vector of observations for vital rate \emph{X}
($X = S, G, R, or P$ for survival, growth, recruitment, or population
growth). For example, $\textbf{y}^S$ is a vector of 0's and 1's
indicating whether a genet survives from \emph{t} to \emph{t+1}, or not.

@@

Our data allows us to track new recruits, but we cannot assign a
specific parent to new genets. THerefore, we model recruitment at the
quadrat level: the number of new individuals of species $j$ in quadrat
$q$ recruiting at time $t+1$ as a function of quadrat ``effective
cover'' ($A'$) in the previous year ($t$). Effective cover is a mixture
of observed cover ($A$) in the focal quadrat ($q$) and the mean cover
across the entire group ($\bar{A}$) of $Q$ quadrats in which $q$ is
located:

@@

The statistical approach used to model aggregated data depends on the
type of data collected. We have percent cover data, which can easily be
transformed to proportion data. We first considered fitting three vital
rate models analagous to those we fit at the individual level: one for
probability of extirpation within a quadrat (analagous to survival), one
for cover change within a quadrat (analagous to growth), and one for
probability of colonization within a quadrat (analagous to recruitment).
However, within-quadrat extirpation and colonization events were rare in
our time series ($N=9$ and $N=10$, respectively, across all species).
Given the broad spatial distribution of the quadrats we are studying, it
is safe to assume that these events are in fact rare enough to be
ignored for our purposes. So we constrained our statistical modeling of
vital rates at the population level to change in percent cover within
quadrats. For the remaining discussion of statistical modeling, we refer
to proportion data, which is simply percent cover divided by 100.

An obvious choice for fitting a linear model to proportion data is beta
regression because the support of the beta distribution is {[}0,1{]},
not including true zeros or ones. However, when we used fitted model
parameters from a beta regression in a quadrat-based population model,
the simulated population tended toward 100\% cover for all species. We
therefore chose a more constrained modeling approach based on a
truncated log-normal likelihood. The model for quadrat cover change from
time $t$ to $t+1$ is

@@

Our Bayesian approach to fitting the vital rate models required choosing
appropriate priors for unknown parameters and deciding which, if any, of
those priors should be hierarchical. We decided to fit models where all
terms were fit by species. Within a species, we fit yearly size effects
and yearly intercepts hierarchically where year-specific coefficients
were drawn from global distributions representing the mean size effect
and intercept. Quadrat random effects were also fit hierarchically, with
quadrat offsets being drawn from distributions with mean zero and a
shared variance term (independent Gaussian priors, Appendix A). Climate
effects were not modeled hierarchically, and each was given a diffuse
prior distribution. We used standard diffuse priors for all unknown
parameters (Appendix A).

All of our analyses (model fitting and simulating) were conducted in R
(R Core Development Team 2013). We used the `No-U-Turn' MCMC sampler in
Stan (Stan Development Team 2014a) to estimate the posterior
distributions of model parameters using the package `rstan' (Stan
Development Team 2014b). We obtained posterior distributions for all
model parameters from three parallel MCMC chains run for 1,000
iterations after discarding an initial 1,000 iterations. Such short MCMC
chains may surprise readers more familiar with other MCMC samplers
(i.e.~JAGS or WinBUGS), but the Stan sampler is exceptionally efficient,
which reduces the number of iterations needed to achieve convergence. We
assessed convergence visually and made sure scale reduction factors for
all parameters were less than 1.01. For the purposes of including
parameter uncertainty in our population models, we saved the final 1,000
iterations from each of the three MCMC chains to be used as randomly
drawn values during population simulation. This step alleviates the need
to reduce model parameters by model selection since sampling from the
full parameter space in the MCMC ensures that if a parameter broadly
overlaps zero, on average the effect in the population models will also
be near zero. We report the posterior mean, standard deviation, and 95\%
Bayesian Credible Intervals for every parameter of each model for each
species in Appendix B.

@@

With the posterior distribution of the vital rate statistical models in
hand, it is straightforward to simulate the population models. We used
an Integral Projection Model (IPM) to model populations based on
individual-level data (Ellner and Rees 2006) and a quadrat-based version
of an individually-based model (Quadrat-Based Model, QBM) to model
populations based on quadrat-level data. We describe each in turn.

@@

We use a stochastic IPM (Rees and Ellner 2009) that includes the climate
covariates from the vital rate statistical models. In all simulations we
ignore the random year effects so that interannual variation is driven
solely by climate. We fit the random year effects in the vital rate
regressions to avoid over-attributing variation to climate covariates.
Our IPM follows the specification of Chu and Adler (2015) where the
population of species \emph{j} is a density function $n(u_{j},t)$ giving
the density of sized-\emph{u} genets at time \emph{t}. Genet size is on
the natural log scale, so that $n(u_{j},t)du$ is the number of genets
whose area (on the arithmetic scale) is between $e^{u_{j}}$ and
$e^{u_{j}+du}$. The density function for any size \emph{v} at time $t+1$
is

\begin{equation}
n(v_{j},t+1) = \int_{L_{j}}^{U_{j}} k_{j}(v_{j},u_{j},\bar{w_{j}}(u_{j}))n(u_{j},t)
\end{equation}

where $k_{j}(v_{j},u_{j},\bar{w_{j}})$ is the population kernel that
describes all possible transitions from size $u$ to $v$ and
$\bar{w_{j}}$ is a scalar representing the average intraspecific
crowding experienced by a genet of size $u_j$ and species $j$. The
integral is evaluated over all possible sizes between predefined lower
(\emph{L}) and upper (\emph{U}) size limits that extend beyond the range
of observed genet sizes.

Since the IPM is spatially-implicit, we cannot calculate neighborhood
crowding for specific genets ($w_{ij}$). Instead, we use an
approximation ($\bar{w_{j}}$) that captures the essential features of
neighborhood interactions (Adler et al. 2010). This approximation relies
on a `no-overlap' rule for conspecific genets to approximate the
overdispersion of large genets in space (Adler et al. 2010).

@@

\begin{equation}
k_{j}(v_{j},u_{j},\bar{w_{j}}) = S_j(u_j, \bar{w_{j}}(u_{j}))G_j(v_{j},u_{j},\bar{w_{j}}(u_{j})) + R_j(v_{j},u_{j},\bar{w_{j}}),
\end{equation}

which means we are calculating growth (\emph{G}) for individuals that
survive (\emph{S}) from time \emph{t} to \emph{t+1} and adding in newly
recruited (\emph{R}) individuals of an average sized one-year-old genet
for the focal species. Our stastical model for recruitment (\emph{R},
described above) returns the number of new recruit produced per quadrat.
Following previous work (Adler et al. 2012, Chu and Adler 2015), we
assume that fecundity increases linearly with size
($R_j(v_{j},u_{j},\bar{w_{j}}) = e^{u_j}R_j(v_{j},\bar{w_{j}})$) to
incorporate the recruitment function in the spatially-implicit IPM.

We used random draws from the final 1,000 iterations from each of three
MCMC chains to introduce stochasticity into our population models. At
each time step, we randomly selected climate covariates from one of the
14 observed years. Then, we drew the full parameter set (climate effects
and density-dependence fixed effects) from a randomly selected MCMC
iteration. Using this approach, rather than simply using coefficient
point estimates, captures the effect of parameter uncertainty.
Relatively unimportant climate covariates (those that broadly overlap 0)
will have little effect on the mean of the simulation results, but can
contribute to their variation. Since our focus was on the contribution
of climate covariates to population states, we set the random year
effects and the random group effects to zero.

@@

To simulate our quad-based model (QBM), we simply iterate the
quadrat-level statistical model (Eqs. 9-10). We use the same approach
for drawing parameter values as described for the IPM. After drawing the
appropriate parameter set, we calculate the mean response (population
cover at \emph{t+1} = $x_{t+1}$) according to Eq. 9. We then make a
random draw from a {[}0,1{]} truncated lognormal distribution with mean
equal to $x_{t+1}$ from Eq. 9 and the variance estimate from the fitted
model. We can then project the model forward by drawing a new parameter
set (unique to climate year and MCMC iteration) at each timestep. As
with the IPM, random year effects are ignored for all simulations.

@@

To test each model's ability to forecast population state, we made
out-of-sample predictions using leave-one-year-out cross validation. For
both levels of modeling, we fit the vital rate models using observations
from all years except one, and then used those fitted parameters in the
population models to perform a one-step-ahead forecast for the year
whose observations were withheld from model fitting. Within each
observation year, several quadrats were sampled. We made predictions for
each observed quadrat in the focal year, initializing each simulation
with cover in the quadrat the previous year. Since we were making
quadrat-specific predictions, we incorporated the group random effect on
the intercept for both models. We repeated this procedure for all 13
observation years, making 100 one-step-ahead forecasts for each
quadrat-year combination with parameter uncertainty included via random
draw from the MCMC chain as described above. Random year effects were
set to zero since year effects cannot be assigned to unobserved years.

This cross-validation procedure allowed us to compare accuracy and
precision of the two modeling approaches (IPM versus QBM). We first
calculated the median predicted cover across the 100 simulations for
each quadrat-year and then calculated the absolute error as the absolute
value of the difference between the observed cover for a given
quadrat-year and the median prediction. To arrive at mean absolute error
(MAE), we then averaged the absolute error within each species across
the quadrat-year specific errors. We use MAE as our measure of accuracy.
To measure precision we calculated the distance between the upper and
lower 90th quantiles of the 100 predictions and averaged this value over
quadrat-years for each species.

@@

We expected the IPM to produce more accurate and precise forecasts due
to either (1) the smaller sample size of the quadrat level data sets
compared to the individual level data sets, leading to larger parameter
uncertainty for the QBM, or (2) the QBM climate effects being weakly
associated with one or more vital rate climate effects at the individual
level. To assess the impact of sample size on QBM parameter uncertainty
we refit the QBM statistical model (Eqs. 9-10) after removing sets of 2,
5, 10, and 15 quadrats. We fit 10 models at each level of quadrat
removal (2, 5, 10, 15 quadrats), removing a different randomly selected
set of quadrats for each fit. We calculated the standard deviation of
climate main effects (pptLag, ppt1, ppt2, TmeanSpr1, and TmeanSpr2) for
each model and averaged those over replicates within each set of quadrat
removals. This allowed us to regress parameter uncertainty against
sample size.

To deterime if the QBM climate effects are correlated with climate
effects for each vital rate model in the IPM, we simply regressed the
QBM climate coefficients against each vital rate model's climate
coefficients and calculated Pearson's $\rho$. Strong correlations
indicate the QBM is capable of detecting climate effects associated with
individual vital rates. A weak correlation indicates the QBM ``misses''
the climate effect on a particular vital rate.

@@
